### 2019-06-17 initial/notes.tex
update notes w/ algebra

\subsection{Separatrix Area}

@@

We notate $\mu = \cos\theta$; note that CS4 is $\mu_4 \approx \frac{\eta \cos
I}{1 - \eta \sin I} \approx \eta \cos I$. Setting the Hamiltonian equal to its
value at CS4 gives
\begin{align*}
    H_4 &\equiv H\p*{\mu_4, \phi_4}
        \approx -\frac{\mu_4^2}{2} + \eta \mu_4 \cos I - \eta \sin I,\\
        &= +\eta^2\cos^2 I - \eta \sin I,\\
    H(\mu_{sep}, \phi_{sep})
        &= H_4 = -\eta \sin I \cos \phi_{sep} - \frac{\mu_{sep}^2}{2} + \eta
            \mu_{sep} \cos I + \mathcal{O}(\eta^3),\\
    0 &\approx \frac{\mu_{sep}^2}{2} - \eta \mu_{sep} \cos I
        - \eta \sin I\p*{1 - \cos \phi_{sep}} + \eta^2\cos^2I,\\
    \mu_{sep}\p*{\phi} &\approx \sqrt{2\eta \sin I\p*{1 - \cos \phi}}
        + \mathcal{O}(\eta).
\end{align*}

@@

\section{Parametric Separatrix Hopping: Only changing $\eta$}

@@

In order to probe the relevant physics here, we firsst cite without proof (refer
to \autoref{ss:revisited} for this) that $\pd{\Delta H}{\alpha} =
\frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I$. This implies that
\begin{equation}
    B_i = \int_{\mathcal{C}_i} \dot{\alpha}\s*{
        \frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I}
        \;\mathrm{d}t.
\end{equation}
Recall that the capture probability is $\frac{B_1 + B_2}{B_1}$ when this
expression is $\in [0, 1]$. However, we immediately see a problem: the first
term will dominate the second term, since $\alpha \gg 1$, but the first term is
also positive semidefinite. This implies $B_1, B_2$ will always have the same
sign unless $\dot{\alpha}$ has different signs over the two trajectories!

@@

First, let's examine our $t_{\star}$ plot e.g.\ \autoref{fig:4_cross0_20}. We
observe that there is a roughly linear slope $\rd{t_{\star}}{(\Delta H)} \sim
\frac{600}{s_0/2s_c}$ where $-s_0/2s_c$ is the minimum value of $H$ on the
domain and so is the maximum value for $\Delta H = H_4 - H_0$. Note that
\begin{align*}
    H(\mu, \phi; s) &= -\frac{s}{s_c}\frac{\mu^2}{2}
        + \mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi, &
    H_4(s) &= -\sin I + \frac{s_c}{2s}\cos^2 I.
\end{align*}
While $s$ is actually a dynamical variable, it is a parameter to the 1D
Hamiltonian. Plugging these expressions into our above expression for $\rd{K}{t}
= -\rd{(\Delta H)}{t}$, we obtain
\begin{align}
    \rd{(\Delta H)}{t} &= -\dot{\mu}^{(1)}\dot{\phi}^{(0)} - \dot{s}^{(1)}
            \pd{(\Delta H)}{s},\\
    -\rd{(\Delta H)}{(\epsilon t)} &= (1 - \mu^2)\p*{\frac{2\Omega}{s} - \mu}
            \dot{\phi}^{(0)} + 2\Omega\p*{1 + \frac{s}{2\Omega}(1 + \mu^2)}
            \s*{\frac{\mu^2}{2s_c} - \frac{s_c}{2s^2}\cos^2 I}.
\end{align}
Note that this expression is encouraging since both $\dot{\phi}^{(0)},
\pd{\p*{\Delta H}}{s}$ vanish near CS4 (where $\mu_4 \approx \frac{s_c}{s}\cos
I$), which corresponds to separatrix approach slowing down (also consistent w/
our plots). Now, if we evaluate far from the separatrix and for large $s$,
similar to most of our initial conditions, we approximate $\dot{\phi} \approx
\frac{s_c}{s}$ and we obtain
\begin{align}
    \rd{(\Delta H)}{(\epsilon t)} &\approx
            \p*{1 - \mu^2}\p*{\frac{2\Omega}{s} - \mu}\frac{s}{s_c}\mu
                + s\p*{1 + \mu^2}\p*{\frac{\mu^2}{2s_c}},\nonumber\\
        &\approx -\frac{s\mu^2}{2s_c} \p*{2(1 - \mu^2) - 1 + \mu^2}
            ,\nonumber\\
        &= -\frac{s\mu^2}{2s_c}\p*{3 - \mu^2}.
\end{align}
The linear fit by eye on the plot corresponds to a slope of $-\rd{(\Delta H)}{t}
\approx \frac{s_0/s_c}{1200}$ which is of the right order $-\epsilon
\frac{3s\mu^2}{2s_c}$, so it seems like we at least have the scaling correct.
The intercept offset is given by the dynamics near $\Delta H$ small, or
sufficiently small that $\dot{\phi}^{(0)} \approx \frac{s}{s_c}$ is no longer
accurate (where phase space begins to be distorted by the separatrix); we will
analyze this separately later/never.

>>>>

\subsection{Separatrix Area}\label{ss:sep_area}

@@

We notate $\mu = \cos\theta$; note that CS4 is $\mu_4 \approx \frac{\eta \cos
I}{1 - \eta \sin I} \approx \eta \cos I$. Setting the Hamiltonian equal to its
value at CS4 gives
\begin{align*}
    H_4 &\equiv H\p*{\mu_4, \phi_4}
        \approx -\frac{\mu_4^2}{2} + \eta \mu_4 \cos I - \eta \sin I,\\
        &= +\frac{\eta^2\cos^2 I}{2} - \eta \sin I,\\
    H(\mu_{sep}, \phi_{sep})
        &= H_4 = -\eta \sin I \cos \phi_{sep} - \frac{\mu_{sep}^2}{2} + \eta
            \mu_{sep} \cos I + \mathcal{O}(\eta^3),\\
    0 &\approx \frac{\mu_{sep}^2}{2} - \eta \mu_{sep} \cos I
        - \eta \sin I\p*{1 - \cos \phi_{sep}} + \frac{\eta^2\cos^2I}{2},\\
    \mu_{sep}\p*{\phi} &\approx \sqrt{2\eta \sin I\p*{1 - \cos \phi}}
        + \mathcal{O}(\eta).
\end{align*}

@@

\section{Parametric Separatrix Hopping: Only changing $\eta$}\label{s:p2}

@@

In order to probe the relevant physics here, we first cite without proof (refer
to \autoref{ss:revisited} for this) that $\pd{\Delta H}{\alpha} =
\frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I$. This implies that
\begin{equation}
    B_i = \int_{\mathcal{C}_i} \dot{\alpha}\s*{
        \frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I}
        \;\mathrm{d}t.
\end{equation}
Recall that the capture probability is $\frac{B_1 + B_2}{B_1}$ when this
expression is $\in [0, 1]$. However, we immediately see a problem: the first
term will dominate the second term, since $\alpha \gg 1$, but the first term is
also positive semidefinite. This implies $B_1, B_2$ will always have the same
sign unless $\dot{\alpha}$ has different signs over the two trajectories!

@@

First, let's examine our $t_{\star}$ plot e.g.\ \autoref{fig:4_cross0_20}. We
observe that there is a roughly linear slope $\rd{t_{\star}}{(\Delta H)} \sim
\frac{600}{s_0/2s_c}$ where $-s_0/2s_c$ is the minimum value of $H$ on the
domain and so is the maximum value for $\Delta H = H_4 - H_0$. Note that
\begin{align*}
    H(\mu, \phi; s) &= -\frac{s}{s_c}\frac{\mu^2}{2}
        + \mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi,&
    H_4(s) &= -\sin I + \frac{s_c}{2s}\cos^2 I.
\end{align*}

I accidentally did some extra algebra and computed to one extra order in
$\frac{s_c}{s}$, so I might as well leave it in. We will compute carefully,
noting that $\mu_4 = \frac{\cos I}{\alpha - \sin I}, \phi_4 = 0$ is exact (where
$\alpha = s/s_c$ for brevity and compatibility with \autoref{s:p2}).
\begin{align*}
    H_4 &= -\alpha \frac{\mu_4^2}{2} + \mu_4 \cos I - \sin I \sqrt{1 - \mu_4^2}
            \cos \phi_4,\\
        &= -\frac{\alpha \cos^2 I}{2(\alpha - \sin I)^2}
            + \frac{\cos^2 I}{\alpha - \sin I}
            - \sin I \frac{\sqrt{(\alpha - \sin I)^2 - \cos^2 I}}{
                \alpha - \sin I},\\
        &= \frac{1}{\p*{\alpha - \sin I}^2}
            \s*{-\alpha \cos^2 I + \p*{\alpha - \sin I}
                \p*{\cos^2 I - \sin I \sqrt{
                    (\alpha - \sin I)^2 - \cos^2 I}}},\\
        &\approx \frac{1}{\p*{\alpha - \sin I}^2}
            \s*{\frac{\alpha \cos^2 I}{2} -\sin I \cos^2 I - \alpha^2 \sin I
                \p*{1 - \frac{2\alpha \sin I + \cos^2 I - \sin^2 I}{2\alpha^2}}
                + \sin^2 I\alpha\p*{1 - \frac{2\alpha \sin I}{2\alpha^2}}},\\
        &\approx \frac{1}{\alpha^2}\p*{1 + \frac{2\sin I}{\alpha}
            + \frac{3\sin^2 I}{\alpha^2}}\p*{
                -\alpha^2 \sin I + 2\sin^2 I\alpha
                    + \frac{\alpha \cos^2 I}{2}
                    - \frac{(\cos^2 I + 3\sin^2 I)\sin I}{2}},\\
        &\approx -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{1}{\alpha^2}\s*{
                -3\sin^3 I - \frac{\p*{\cos^2 I + 3\sin^2 I}\sin I}{2}
                    + 4\sin^3 I + \sin I \cos^2 I},\\
        &= -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{\p*{\cos^2 I - \sin^2 I}\sin I}{2\alpha^2}.
\end{align*}

While $s$ is actually a dynamical variable, it is a parameter to the 1D
Hamiltonian. Plugging these expressions into our above expression for $\rd{K}{t}
= -\rd{(\Delta H)}{t}$, we obtain
\begin{align}
    \rd{(\Delta H)}{t} &= -\dot{\mu}^{(1)}\dot{\phi}^{(0)} - \dot{s}^{(1)}
            \pd{(\Delta H)}{s},\\
    -\rd{(\Delta H)}{(\epsilon t)} &= (1 - \mu^2)\p*{\frac{2\Omega}{s} - \mu}
            \dot{\phi}^{(0)} + 2\Omega\p*{1 + \frac{s}{2\Omega}(1 + \mu^2)}
            \s*{\frac{\mu^2}{2s_c} - \frac{s_c}{2s^2}\cos^2 I}.
\end{align}
Note that this expression is encouraging since both $\dot{\phi}^{(0)},
\pd{\p*{\Delta H}}{s}$ vanish near CS4 (where $\mu_4 \approx \frac{s_c}{s}\cos
I$), which corresponds to separatrix approach slowing down (also consistent w/
our plots). Now, if we evaluate far from the separatrix and for large $s$,
similar to most of our initial conditions, we approximate $\dot{\phi} \approx
\frac{s_c}{s}$ and we obtain
\begin{align}
    \rd{(\Delta H)}{(\epsilon t)} &\approx
            \p*{1 - \mu^2}\p*{\frac{2\Omega}{s} - \mu}\frac{s}{s_c}\mu
                + s\p*{1 + \mu^2}\p*{\frac{\mu^2}{2s_c}},\nonumber\\
        &\approx -\frac{s\mu^2}{2s_c} \p*{2(1 - \mu^2) - 1 + \mu^2}
            ,\nonumber\\
        &= -\frac{s\mu^2}{2s_c}\p*{3 - \mu^2}.
\end{align}
The linear fit by eye on the plot corresponds to a slope of $-\rd{(\Delta H)}{t}
\approx \frac{s_0/s_c}{1200}$ which is of the right order $-\epsilon
\frac{3s\mu^2}{2s_c}$, so it seems like we at least have the scaling correct.
The intercept offset is given by the dynamics near $\Delta H$ small, or
sufficiently small that $\dot{\phi}^{(0)} \approx \frac{s}{s_c}$ is no longer
accurate (where phase space begins to be distorted by the separatrix); we will
analyze this separately later/never.
